Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-01-28 colt-21/cvx-flats.tex
comment
Reading over Raf's comments

% vertices
\node[vertex] (consis-ell) at  (-4,3) {Consistency w.r.t. $\ell$};
\node[vertex] (consis-gam) at  (2,3) {Consistency w.r.t. $\gamma$};
\node[vertex] (indir-elic) at  (-4,-3) {Indirect elicitation};
\node[vertex] (d-flats) at  (2,-3) {$d$-flats};
\node[vertex] (conscvx-bd) at (8,0) {$\conscvx(\gamma) > d$};
\node[vertex] (q1-bounds) at (14,3) {$\conscvx$ bounds for Q1};
\node[vertex] (q4-bounds) at (14,-3) {$\conscvx$ bounds for Q4};

@@

The \emph{level set} of $\Gamma$ at value $r\in\R$ is $\Gamma_r := \{p \in \P : r \in \Gamma(p)\}$.
We call a property $\Gamma: \P \toto \R$ \emph{discrete} if $\R$ is a finite set.
A property is \emph{single-valued} if $|\Gamma(p)| = 1$ for all $p\in\P$.
A property is \emph{single-valued} if $|\Gamma(p)|=1$ for all $p\in\P$, in which case we may write $\Gamma:\P\to\R$ and $\Gamma(p) \in \R$\raft{added, since sec 6 gets confusing without this single-arrow notation}.
We define the \emph{range} of a property by $\range \Gamma = \bigcup_{p\in\P} \Gamma(p) \subseteq \R$.
When $L\in\L$, we use $\Gamma := \prop[\P]{L}$ to denote the unique property elicited by $L$ (for distributions in $\P$) from eq.~\eqref{eq:elic}. 
Typically, we denote the target property by $\gamma$, and the surrogate by $\Gamma$.

@@

An important caveat to the above definitions is that, since $\Gamma = \prop[\P]{L}$ is nonempty everywhere, we must have $L\in\L$, meaning that $\exploss{L}{\cdot}{p}$ always achieves a minimum.
This restriction is also implicit in e.g.~\citep{agarwal2015consistent}.
While some popular surrogates such as logistic and exponential loss are not minimizable, these losses are still covered in Corollary~\ref{cor:Pcodim-flat-elic-relint-prop} and Theorem~\ref{thm:bayes-risk-lower-bound} as $\Gamma(p) \neq \emptyset$ when $p\in\P := \relint\simplex$; moreover, by thresholding $L'(u,y) = \max(L(u,y),\epsilon)$ for sufficiently small $\epsilon>0$ we can achieve $L'\in\L$ for both.
We expect that a generalization of property elicitation which allows for ``infinite'' predictions (e.g., along a prescribed ray), thereby ensuring a minimum is always achieved for convex losses, would allow us to lift the minimizable restriction entirely.
\raf{Check me!}
\btw{Some refs / related work commented out here.  I think not as relevant as elicitation complexity.}

@@

\begin{definition}[Convex Consistency Dimension]\label{def:cvx-consistency-dim}
  Given a target loss $\ell:\R \times\Y \to \reals$ or statistic $\gamma: \P \toto \R$, its \emph{convex consistency dimension} $\conscvx(\cdot)$ is the minimum dimension $d$ such that there is a convex $L : \reals^d \times \Y \to \reals$ and link $\psi$ such that $(L,\psi)$ is consistent with respect to $\ell$ or $\gamma$.
  \raft{Why not use $\Lcvx$ here?  Then you don't need the qualifier below.}
%	\btw{Maybe take out of definition environment, but (a) I'd worry that people wouldn't be able to find the definition when we use it later, and (b) I'm hoping we can make the distinction very clear, which now is achieved by the parallel definitions}
\end{definition}
A priori, this definition does not really make sense given a target statistic.
In Definition~\ref{def:consistent-prop}, we formalize consistency with respect to a property so that we can study $\conscvx(\gamma)$. \jessiet{Do we want to just talk about $\ell$ here?}\raft{I didn't think about the target property vs target loss point.  Let's discuss.}

@@

\begin{definition}[Consistent: loss]\label{def:consistent-ell}
  A loss and link $(L,\psi)$ are \emph{$\D$-consistent} for a set $\D$ of distributions over $\X \times \Y$ with respect to a target loss $\ell$ if, for all $D \in \D$ and all sequences of measurable hypothesis functions $\{f_m : \X \to \R\}$,
  \begin{align*}
    \E_D L(f_m(X), Y) \to \inf_f \E_D L(f(X), Y) &\implies \E_D \ell((\psi \circ f_m)(X), Y) \to \inf_f \E_D \ell((\psi \circ f)(X), Y)~.~
  \end{align*}
  % Given a set $\P \subseteq \simplex$, if for all distributions $p \in \P$, there exists a $D \in \D$ and $x \in \X$ such that $D$ has a point mass on $x$ and $p = D_x$, then we simply say $(L,\psi)$ is consistent.
  For a given set $\P \subseteq \simplex$, we simply say $(L,\psi)$ is \emph{consistent} if it is $\D$-consistent for some $\D$ satisfying the following: for all $p \in \P$, there exists $D \in \D$ and $x \in \X$ such that $D$ has a point mass on $x$ and $p = D_x$.
  \raft{Touched up to clarify the quantification over $\D$}
\end{definition}

@@

\begin{definition}[Consistent: property]\label{def:consistent-prop}
	Suppose we are given a loss $L \in \L_d$, link function $\psi: \reals^d \to \R$, and property $\gamma:\P \toto \R$.
	Moreover, let $\propdis : \R \times \P \to \reals_+$ be any function satisfying $\propdis(r,p) = 0 \iff r \in \gamma(p)$.
	We say $(L, \psi)$ is \emph{$(\propdis, \D)$-consistent with respect to} $\gamma$ if, for all $D \in \D$ and sequences of measurable functions $\{f_m: \X \to \R\}$, 
	\begin{equation}
    \E_{D} L(f_m(X), Y) \to \inf_f \E_{D} L( f(X), Y) \implies \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~
  \end{equation}
  % where $D_x = \Pr[Y|X=x]$.
  We simply say $(L,\psi)$ is \emph{$\mu$-consistent} if it is $(\mu,\D)$-consistent for some $\D$ satisfying the following: for all $p \in \P$, there exists $D \in \D$ and $x \in \X$ such that $D$ has a point mass on $x$ and $p = D_x$.
  \raft{Ditto; hopefully this is equivalent to what was here before (commented out below)}
  % Suppose there is a set $\D$ such that, for all $p \in \P$, there is a $D \in \D$ with a point mass on some $x \in \X$ such that $p = D_x$ and a $\propdis$ satisfying the requirements above.
  % In this case, if $(L,\psi)$ are $(\propdis,\D)$-consistent, we simply say the pair is consistent with respect to $\gamma$.
\end{definition}

@@

\begin{proof}
  By Lemma~\ref{lem:consistent-loss-implies-prop}, it suffices to show the result for consistency with respect to a property $\gamma$, setting $\gamma := \prop{\ell}$ if $\ell$ is given instead.
  We show the contrapositive; suppose $(L, \psi)$ does not indirectly elicit $\gamma$, meaning we have some $p \in \P$ so that $u \in \Gamma(p)$ but $\psi(u) \not \in \gamma(p)$, where $\Gamma := \prop{L}$.
  Observe that we use the fact $\Gamma(p) \neq \emptyset$.
  By definition, we have consistency with respect to some distribution $D$ on $\X\times\Y$ with a point mass on some $x\in\X$ and $D_x = p$.
  Consider a constant sequence $\{f_m\}$ with $f_m = f'$ such that $f'(x) = u$,
  so that $\E_D L(f_m(X), Y) = \E_{D_x} L(f_m(x),Y) = \E_p L(u,Y)$.
  Since $u \in \Gamma(p)$, we have $\E_p L(u,Y) = \inf_f \E_{D_x} L(f(x),Y) = \inf_f \E_D L(f(X),Y)$.
  In particular, we have $\E_D L(f_m(X), Y) \to \inf_f \E_D L(f(X),Y)$.
  \raft{Old version commented out; I think we meant $D$ not $p$}%$\E_p L(f_m(x), Y) \to \inf_f \E_p L(f(X),Y)$.
  % \raft{Maybe clearer to first say $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$, and then the easy implication of the convergence \jessie{Done}}
  However, we have $\E_X \propdis(\psi \circ f_m(X), D_X) = \propdis(\psi(u_m), p) = \propdis(\psi(u), p) \neq 0$, since $\psi(u) \not \in \gamma(p)$.
  \raft{$u_m$ not defined; just cut out that term perhaps?}
  Therefore $(L, \psi)$ is not consistent with respect to $\gamma$ (Definition~\ref{def:consistent-prop}).
\end{proof}

@@

We now state our elicitation lower bounds in Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}, which when combined with Theorem \ref{thm:consistent-implies-indir-elic}, implies consistency bounds.
A similar result is \citet[Theorem 9]{agarwal2015consistent}, which bounds the dimension of level sets of a single-valued $\prop{L}$.
Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop} instead bound the dimension of flats contained in the level sets, an additional power which we leverage in our examples.
%Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop} also give a lower bound for any given target $\ell$ or $\gamma$, rather than a given surrogate $L$. \jessie{Does AA15 do this?  Seems out of the blue without that context.}
\raf{Cut the last bits here; I bet AA15 thought of their result as being about consistency, though I could be wrong}
% (although an analogous implication of their Theorem 9 may have been known to \citet{agarwal2015consistent}).
% We also note that the lower bound on elicitation complexity in \citet{agarwal2015consistent} is not translated into a lower bound for \emph{consistent} surrogates, which is the main contribution of this work.

@@

\begin{lemma}\label{lem:finite-relint-dim}
  Let the $d$-flat $F\subseteq \P$ (defined over finite $\Y$) contain some $p\in\relint{\P}$.
  Then 
  \begin{enumerate}
  	\item[(i)] $p \in \relint{F}$; 
  	\item[(ii)] $\dim(\Scr_F(p)) \geq \dim(\affhull(\P)) - d$.
  \end{enumerate}
\end{lemma}
\jessiet{Bo had a good point that $\spn(F_p)$ is probably easier to understand, instead of having an extra layer of notation in there.  Do we want to stick with $\Scr_F(p)$ or $\spn(F_p)$?}
\begin{proof}
  %\raf{Moved; see margin comment:  for the function $W : \Y \to \reals^d$ such that $F = \zeros{W}$.  (Such a function must exist by $F$ being a $d$-flat.)}

>>>>

% vertices
\node[vertex] (consis-ell) at  (-4,3) {Consistency w.r.t. $\ell$};
\node[vertex] (consis-gam) at  (2,3) {Consistency w.r.t. $\gamma$};
\node[vertex] (indir-elic) at  (-4,-3) {Indirect elicitation};
\node[vertex] (d-flats) at  (2,-3) {$d$-flats};
\node[vertex] (conscvx-bd) at (8,0) {$\conscvx(\gamma) \leq d$};
\node[vertex] (q1-bounds) at (14,3) {$\conscvx$ bounds for Q1};
\node[vertex] (q4-bounds) at (14,-3) {$\conscvx$ bounds for Q4};

@@

The \emph{level set} of $\Gamma$ at value $r\in\R$ is $\Gamma_r := \{p \in \P : r \in \Gamma(p)\}$.
We call a property $\Gamma: \P \toto \R$ \emph{discrete} if $\R$ is a finite set.
A property is \emph{single-valued} if $|\Gamma(p)| = 1$ for all $p\in\P$.
A property is \emph{single-valued} if $|\Gamma(p)|=1$ for all $p\in\P$, in which case we may write $\Gamma:\P\to\R$ and $\Gamma(p) \in \R$.
We define the \emph{range} of a property by $\range \Gamma = \bigcup_{p\in\P} \Gamma(p) \subseteq \R$.
When $L\in\L$, we use $\Gamma := \prop[\P]{L}$ to denote the unique property elicited by $L$ (for distributions in $\P$) from eq.~\eqref{eq:elic}. 
Typically, we denote the target property by $\gamma$, and the surrogate by $\Gamma$.

@@

An important caveat to the above definitions is that, since $\Gamma = \prop[\P]{L}$ is nonempty everywhere, we must have $L\in\L$, meaning that $\exploss{L}{\cdot}{p}$ always achieves a minimum.
This restriction is also implicit in e.g.~\citep{agarwal2015consistent}.
While some popular surrogates such as logistic and exponential loss are not minimizable, these losses are still covered in Corollary~\ref{cor:Pcodim-flat-elic-relint-prop} and Theorem~\ref{thm:bayes-risk-lower-bound} as $\Gamma(p) \neq \emptyset$ when $p\in\P := \relint\simplex$; moreover, by thresholding $L'(u,y) = \max(L(u,y),\epsilon)$ for sufficiently small $\epsilon>0$ we can achieve $L'\in\L$ for both.
We expect that a generalization of property elicitation which allows for ``infinite'' predictions (e.g., along a prescribed ray), thereby ensuring a minimum is always achieved for convex losses, would allow us to lift the minimizable restriction entirely.
\raf{Check me!}\jessie{<< From NeurIPS.  I'm good with this}
\btw{Some refs / related work commented out here.  I think not as relevant as elicitation complexity.}

@@

\begin{definition}[Convex Consistency Dimension]\label{def:cvx-consistency-dim}
  Given a target loss $\ell:\R \times\Y \to \reals$ or statistic $\gamma: \P \toto \R$, its \emph{convex consistency dimension} $\conscvx(\cdot)$ is the minimum dimension $d$ such that there is a convex $L :\reals^d \times\Y \to \reals$ and link $\psi$ such that $(L,\psi)$ is consistent with respect to $\ell$ or $\gamma$.
  \raft{Why not use $\Lcvx$ here?  Then you don't need the qualifier below. \jessie{$\Lcvx$ implies minimizability, do we want that baked in here?}}
%	\btw{Maybe take out of definition environment, but (a) I'd worry that people wouldn't be able to find the definition when we use it later, and (b) I'm hoping we can make the distinction very clear, which now is achieved by the parallel definitions}
\end{definition}
A priori, this definition does not really make sense given a target statistic.
In Definition~\ref{def:consistent-prop}, we formalize consistency with respect to a property so that we can study $\conscvx(\gamma)$. \jessiet{Do we want to just talk about $\ell$ here?}\raft{I didn't think about the target property vs target loss point.  Let's discuss.}

@@

\begin{definition}[Consistent: loss]\label{def:consistent-ell}
  A loss and link $(L,\psi)$ are \emph{$\D$-consistent} for a set $\D$ of distributions over $\X \times \Y$ with respect to a target loss $\ell$ if, for all $D \in \D$ and all sequences of measurable hypothesis functions $\{f_m : \X \to \R\}$,
  \begin{align*}
    \E_D L(f_m(X), Y) \to \inf_f \E_D L(f(X), Y) &\implies \E_D \ell((\psi \circ f_m)(X), Y) \to \inf_f \E_D \ell((\psi \circ f)(X), Y)~.~
  \end{align*}
  % Given a set $\P \subseteq \simplex$, if for all distributions $p \in \P$, there exists a $D \in \D$ and $x \in \X$ such that $D$ has a point mass on $x$ and $p = D_x$, then we simply say $(L,\psi)$ is consistent.
  For a given set $\P \subseteq \simplex$, we simply say $(L,\psi)$ is \emph{consistent} if it is $\D$-consistent for some $\D$ satisfying the following: for all $p \in \P$, there exists $D \in \D$ and $x \in \X$ such that $D$ has a point mass on $x$ and $p = D_x$.
  %\raft{Touched up to clarify the quantification over $\D$ \jessie{What I was trying to say, but better.}}
\end{definition}

@@

\begin{definition}[Consistent: property]\label{def:consistent-prop}
	Suppose we are given a loss $L \in \L_d$, link function $\psi: \reals^d \to \R$, and property $\gamma:\P \toto \R$.
	Moreover, let $\propdis : \R \times \P \to \reals_+$ be any function satisfying $\propdis(r,p) = 0 \iff r \in \gamma(p)$.
	We say $(L, \psi)$ is \emph{$(\propdis, \D)$-consistent with respect to} $\gamma$ if, for all $D \in \D$ and sequences of measurable functions $\{f_m: \X \to \R\}$, 
	\begin{equation}
    \E_{D} L(f_m(X), Y) \to \inf_f \E_{D} L( f(X), Y) \implies \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~
  \end{equation}
  % where $D_x = \Pr[Y|X=x]$.
  We simply say $(L,\psi)$ is \emph{$\propdis$-consistent} if it is $(\propdis,\D)$-consistent for some $\D$ satisfying the following: for all $p \in \P$, there exists $D \in \D$ and $x \in \X$ such that $D$ has a point mass on $x$ and $p = D_x$.
  Additionally, we say $(L,\psi)$ is \emph{consistent} if there is a $\propdis$ such that $(L,\psi)$ is $\propdis$-consistent.
  \raft{Ditto; hopefully this is equivalent to what was here before (commented out below)\jessie{Looks good to me}}
  % Suppose there is a set $\D$ such that, for all $p \in \P$, there is a $D \in \D$ with a point mass on some $x \in \X$ such that $p = D_x$ and a $\propdis$ satisfying the requirements above.
  % In this case, if $(L,\psi)$ are $(\propdis,\D)$-consistent, we simply say the pair is consistent with respect to $\gamma$.
\end{definition}

@@

\begin{proof}
  By Lemma~\ref{lem:consistent-loss-implies-prop}, it suffices to show the result for consistency with respect to a property $\gamma$, setting $\gamma := \prop{\ell}$ if $\ell$ is given instead.
  We show the contrapositive; suppose $(L, \psi)$ does not indirectly elicit $\gamma$, meaning we have some $p \in \P$ so that $u \in \Gamma(p)$ but $\psi(u) \not \in \gamma(p)$, where $\Gamma := \prop{L}$.
  Observe that we use the fact $\Gamma(p) \neq \emptyset$.
  By definition, we have consistency with respect to some distribution $D$ on $\X\times\Y$ with a point mass on some $x\in\X$ and $D_x = p$.
  Consider a constant sequence $\{f_m\}$ with $f_m = f'$ such that $f'(x) = u$,
  so that $\E_D L(f_m(X), Y) = \E_{D_x} L(f_m(x),Y) = \E_p L(u,Y)$.
  Since $u \in \Gamma(p)$, we have $\E_p L(u,Y) = \inf_f \E_{D_x} L(f(x),Y) = \inf_f \E_D L(f(X),Y)$.
  In particular, we have $\E_D L(f_m(X), Y) \to \inf_f \E_D L(f(X),Y)$.
  \raft{Old version commented out; I think we meant $D$ not $p$}%$\E_p L(f_m(x), Y) \to \inf_f \E_p L(f(X),Y)$.
  % \raft{Maybe clearer to first say $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$, and then the easy implication of the convergence \jessie{Done}}
  However, we have $\E_X \propdis(\psi \circ f_m(X), D_X) = \propdis(f_m(x), p) = \propdis(\psi(u), p) \neq 0$, since $\psi(u) \not \in \gamma(p)$.
  \raft{$u_m$ not defined; just cut out that term perhaps?}
  Therefore $(L, \psi)$ is not consistent with respect to $\gamma$ (Definition~\ref{def:consistent-prop}).
\end{proof}

@@

We now state our elicitation lower bounds in Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}, which when combined with Theorem \ref{thm:consistent-implies-indir-elic}, implies consistency bounds.
A similar result is \citet[Theorem 9]{agarwal2015consistent}, which bounds the dimension of level sets of a single-valued $\prop{L}$.
Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop} instead bound the dimension of flats contained in the level sets, an additional power which we leverage in our examples.
%Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop} also give a lower bound for any given target $\ell$ or $\gamma$, rather than a given surrogate $L$. \jessie{Does AA15 do this?  Seems out of the blue without that context.}
\raf{Cut the last bits here; I bet AA15 thought of their result as being about consistency, though I could be wrong}\jessie{Old comment $\sim$ NeurIPS.  Good to burn?} 
% (although an analogous implication of their Theorem 9 may have been known to \citet{agarwal2015consistent}).
% We also note that the lower bound on elicitation complexity in \citet{agarwal2015consistent} is not translated into a lower bound for \emph{consistent} surrogates, which is the main contribution of this work.

@@

\begin{lemma}\label{lem:finite-relint-dim}
  Let the $d$-flat $F\subseteq \P$ (defined over finite $\Y$) contain some $p\in\relint{\P}$.
  Then 
  \begin{enumerate}
  	\item[(i)] $p \in \relint{F}$; 
  	\item[(ii)] $\dim(\Scr_F(p)) \geq \dim(\affhull(\P)) - d$.
  \end{enumerate}
\end{lemma}
\jessiet{Bo had a good point that $\spn(F_p)$ is probably easier to understand, instead of having an extra layer of notation in there.  Do we want to stick with $\Scr_F(p)$ or $\spn(F_p)$?}
\raft{True, but the reader doesn't know $F_p$ anymore, since it's within the proof.  Plus, this is the statement we really need.  So I'd vote sticking with $\Scr_F$.}
\begin{proof}
  %\raf{Moved; see margin comment:  for the function $W : \Y \to \reals^d$ such that $F = \zeros{W}$.  (Such a function must exist by $F$ being a $d$-flat.)}
